---
output:
  pdf_document: default
---
```{r setup, include=FALSE, echo=F}
# rm(list=objects())
# ############
# # 0. Custom code 
# ###########
# name <- "Hua Ling"
# proj <- "testShaffer" # this one needs to be the same as the exising directory where your plate map outputs are located.
# array <-'Illumina Multi Ethnic Global Array'
# 
# dir.input <- "R:/active/Shaffer_DentalCaries_MultiEthnicGlobal_033117_2/Project_Management/Plate_Map_Design/ForMakePlateMap"
#   phen.name <- "Shaffer_DentalCaries_MultiEthnicGlobal_033117_2_SIF_Aliquot_040219_2523.csv"
#   dup.filename <- "DUPS_Aliquot_Report_040219_25534.csv"
#   rearray.filename <- "P2104_Shaffer_REARRAYSENT_2018-05-16-08_46_45.109.csv"
# 
# pm.filename <- "testShaffer_RearrayManifest_2019-04-19.csv" 
# vars_to_stratify <- "NA" # Or put in sth like c("Sex", "Population")
# vars_to_check <- c("Sex", "Population")
# random_seed <- "365"

rm(list=objects())
############
# 0. Custom code 
###########
name <- "Hua Ling"
proj <- "testAdebamowo" # this one needs to be the same as the exising directory where your plate map outputs are located.
array <-'H3African'

dir.input <- "R:/active/Adebamowo_CervicalCa_H3Africa_033117_1/Project_Management/Plate_Maps/Files_for_Genetic_Analysts"
  phen.name <- "Adebamowo_CervicalCa_H3Africa_033117_1_SIF_Aliquot_022619_81243_modified_by_PM_for_GA.csv"
  dup.filename <- "DUPS_Aliquot_Report_022619_7742_modified_by_PM_for_GA.csv"
  rearray.filename <- "P2112_Adebamowo_REARRAYSENT_2019-02-21-13_47_15.916.csv"

pm.filename <- "testAdebamowo_RearrayManifest_2019-05-01.csv" 
vars_to_stratify <- c("MTA") # Or put in sth like c("Sex", "Population")
vars_to_check <- c("Population")
random_seed <- "365"

```

---
title: "`r proj` Plate Map Report"
author: "`r name` at CIDR"
date: "`r format(Sys.time(), '%d %B, %Y')`"
# date: "05/25/2018"
output: pdf_document
header-includes:
   - \usepackage{rotating}
---

```{r, include=F, echo=F}
knitr::opts_chunk$set(echo = FALSE)
library(xtable)
library(dplyr)
dir <- paste0("R:/statgen/UW_PlateMap/",proj)
```


# Overview
```{r, results='hide'}
# read final phenotype file
phen.file <- file.path(dir.input, phen.name)
  phen <- read.csv(phen.file, as.is=TRUE, header=TRUE)

# read dup file 
dup.file <- file.path(dir.input, dup.filename)
  dup <- read.csv(dup.file, as.is=TRUE, header=TRUE)
  dupname <- names(dup)[grep("Subject", names(dup))]
  
# read plate map file
pm.file <- file.path("R:/statgen/UW_PlateMap", proj, "output", pm.filename)
  pm <- read.csv(pm.file, as.is=TRUE, header=TRUE, skip=5)
  pm <- pm[,1:3]
  names(pm)[1] <- "Plate"
  names(pm)[3] <- "Subject_ID"
  nctrl <- length(grep("CONTROL", pm$Subject_ID))
# merge plate map and phenotype file
  pm.phen <- merge(pm, phen, by=c("Subject_ID"))
  
# all variables
  vars_all <- c(vars_to_stratify, vars_to_check)
```


  |  
--- | ---
Stratification variables | None
Study samples | `r sum(pm.phen$DNA_Class == 1)`
Duplicates | `r sum(pm.phen$DNA_Class == 2)`
Cross-study duplciates | `r sum(pm.phen$DNA_Class == 3)`
Controls | `r nctrl`
Array | `r array`
Batch size | 96
Number of plates | `r length(unique(pm$Plate))`


# Files used to make plate map

- phenotype file:
    - `r phen.name`
    - Source: PI and CIDR
    - Variables: `r vars_all`
- dupliate pair file:
    - `r dup.filename`
    - Source: CIDR ProjMng
    - Variables: `r dupname`
- rearray file:
    - `r rearray.filename`
    - Source: CIDR ProjMng (Marcia or Tameka)
- Random seed: `r random_seed`


# Plate Map Design

Samples were assigned to plates at random, with all members of a family placed on the same plate.


# Checks
```{r}
n_cntl_plate <- nctrl/length(unique(pm$Plate))
n_crossStudyDup_plate <- sum(pm.phen$DNA_Class == 3)/length(unique(pm$Plate))
n_dup_plate <- sum(pm.phen$DNA_Class == 2)/length(unique(pm$Plate))
```

- All family members are placed on the same plate.
- None of `r vars_to_check` has a significant association with plate. The tables of plate balance by `r vars_to_check` show all samples, but only one randomly-selected subject per family was used in the calculation of p-value for the distribution by plate.
- Duplicate samples are not placed on the same plate as the original sample.
- Each plate has either `r ceiling(n_dup_plate)` or `r floor(n_dup_plate)` duplicates.
- Cross study duplicates are evenly distributed across plates; each plate has either `r ceiling(n_crossStudyDup_plate)` or `r floor(n_crossStudyDup_plate)` cross-study duplicate samples.
- Each plate has `r n_cntl_plate` HapMap/control samples.

\clearpage

```{r, echo=FALSE, results='asis', warning=F}
one_per_family <- pm.phen %>%
  group_by(Family) %>%
  sample_n(size = 1)

for (var in vars_to_check) {
  x_all <- table(pm.phen$Plate, pm.phen[[var]])
  pval_all <- suppressWarnings(chisq.test(x_all)$p.value)
  char_var <- sum(nchar(names(table(pm.phen[,var]))))
  if (!all(pm.phen$Family==0)){
    x_one <- table(one_per_family$Plate, one_per_family[[var]])
    pval_one <- suppressWarnings(chisq.test(x_one)$p.value)
    caption <- paste(sprintf("Distribution of %s by plate (p = %4.3g).", var, pval_one))
  } else if (all(pm.phen$Family==0)){
    caption <- paste(sprintf("Distribution of %s by plate (p = %4.3g).", var, pval_all))
  }
  label <- sprintf("tab:%s", var)
  if (char_var >=65){
    print(xtable(x_all, caption = caption, label=label), type="latex"
          , rotate.colnames = T
          , comment=FALSE
          , tabular.environment="longtable")
  } else {
    print(xtable(x_all, caption = caption, label=label), type="latex", 
          comment=FALSE
          , tabular.environment="longtable")
  }
   cat("\n\n\\pagebreak\n")
}
```

\clearpage

# Duplicates
```{r, results='asis', warning=FALSE}
dup <- dup[dup$Duplicate.Type == "ISD" & dup$Pair.Status == "Active", dupname]
dup$plate1 <- NA; dup$plate2 <- NA
for(i in 1:nrow(dup)) {
  dup$plate1[i] <- pm$Plate[is.element(pm$Subject_ID,dup[i,1])]
  dup$plate2[i] <- pm$Plate[is.element(pm$Subject_ID,dup[i,2])]
}
print(xtable(dup), include.rownames=FALSE, size = "small", type = "latex"
      , comment=FALSE
      , tabular.environment="longtable")

```
